---
title: "DA5020.A3.Xinhao.Sun"
author: "Xinhao Sun"
date: "9/21/2022"
output: 
  html_document:
    code_folding: "hide"
---

## Question 1

#### I completed the data camp "Introduction to the Tidyverse and Reshaping Data with tidyr". 

## Question 2

```{r}
library(tidyverse)
tripdata_df = read.csv("2018_Green_Taxi_Trip_Data.csv")
dim(tripdata_df)
glimpse(tripdata_df)
summary(tripdata_df)
```

## Question 3
### (a)
```{r}
pa = tripdata_df %>%
  select(passenger_count) %>%
  # filter(VendorID != 1 &VendorID != 2)
  filter(passenger_count<1)

dis = tripdata_df %>%
  select(trip_distance) %>%
  filter(trip_distance<=0)

#tripdata_df %>%
#  filter(store_and_fwd_flag != "Y" & store_and_fwd_flag != "N")

fare = tripdata_df %>%
  select(fare_amount) %>%
  filter(fare_amount < 0)

"no passenger"
glimpse(pa)
"zero distance"
glimpse(dis)
"negative fare on meter"
glimpse(fare)

```

```{r, results="hide"}
# show the result above in one table
tripdata_df %>%
  select(passenger_count, trip_distance, fare_amount) %>%
  filter(passenger_count<1 |
           trip_distance<=0 |
           fare_amount<0)
```

### (b)
```{r}
mod_vendorID = factor(tripdata_df$VendorID,levels=c(1,2),labels=c('Creative Mobile Technologies','VeriFone Inc.'))
"VendorID"
glimpse(mod_vendorID)

mod_store_fwd_flag = factor(tripdata_df$store_and_fwd_flag,levels=c("Y","N"),labels=c('store and forward trip','not a store and forward trip'))
"store_and_fwd_flag"
glimpse(mod_store_fwd_flag)

mod_trip_type = factor(tripdata_df$trip_type,levels=c(1,2),labels=c('Street-hail','Dispatch'))
"trip_type"
glimpse(mod_trip_type)

```


## Question 4
```{r, warning=FALSE}
library(tidyverse)
trip_dis = tripdata_df %>%
  select(trip_distance) %>%
  group_by(trip_distance) %>%
  summarize(n = n())

ggplot(trip_dis, aes(trip_distance)) + 
  geom_histogram(fill='#66CCFF')

ggplot(trip_dis, aes(trip_distance)) + 
  geom_histogram(fill='#66CCFF') +
  scale_x_log10() + scale_y_log10()
```

#### we can easily see it is negative skewed data,after using a log scale, the bars varies dramatically.

## Question 5
```{r, results='hide'}

# first method
mean_tip = mean(tripdata_df$tip_amount)
std_tip = sd(tripdata_df$tip_amount)
newdate = tripdata_df %>%
  select(tip_amount,trip_distance) %>%
  filter(tip_amount > mean_tip + 3 * std_tip |
       tip_amount < mean_tip - 3 * std_tip)

```

```{r }

# using reusble function to get outiler should be more easiy

outliers = function(df,a){
  library(dplyr)
  mean_a = mean(a)
  std_a = sd(a)
  newdata = df %>% filter(a > mean_a + 3 * std_a | a < mean_a - 3 * std_a)
  return(newdata)
}

dis_out=outliers(tripdata_df, tripdata_df$trip_distance)
tip_out=outliers(tripdata_df, tripdata_df$tip_amount)

"there are around 2% outliers in trip_distance"
20953/1048575
glimpse(dis_out)

14444/1048575
"there are around 1% outliers in tip_amount"
glimpse(tip_out)
# newdata = setdiff(tripdata_df, tip_out)
# why the result by using "setdiff" is different with the result below
newdata = tripdata_df %>%
  filter(tip_amount < mean(tip_amount) + 3 * sd(tip_amount) & 
           tip_amount > mean(tip_amount) - 3 * sd(tip_amount))

```
#### there are around 2% outliers in trip_distance and there are around 1% outliers in tip_amount


## Question 6
```{r warning=FALSE}
glimpse(factor(tripdata_df$payment_type))
# payment_type have 5 factors instead of 6 which is shown on dictionary

trip_type = newdata %>%
  group_by(payment_type) %>%
  summarise(trips = n()) %>%
  select(trips) %>%
  mutate(pay_type=c('Credit card','Cash', 'No charge', 'Dispute', 'Unknown'))
ggplot(trip_type, aes(x=pay_type, y=trips, fill=pay_type)) + geom_col() +
  scale_y_log10() +
  labs(x = 'Pay Type', y = 'log of Number', title = 'Payment Type for NYC Taxi')
```



## Question 7

#### As far as I am concerned, there are na.omit(), drop_na() and fill(). I recommend fill(), since the amount of missing values is so large that the analysis results could be influenced.

#### An example examples on Taxi data is shown below, I prefer use drop_na() on trip_type due to only 3 NA
```{r}
nonna_type = tripdata_df %>%
  select(trip_type) %>%
  drop_na()
dim(tripdata_df)
dim(nonna_type)
```


