---
title: "practicum2.Xinhao.Sun"
author: "Xinhao Sun"
date: "11/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## initial
```{r, message=FALSE}
rm(list=ls())
library(RSQLite)
library(tidyverse)
library(ggplot2)
library(knitr)
library(kableExtra)
```

# Part 1

## question 1
```{r}
my_database = src_sqlite("practicum2_p1.db", create = TRUE) # create =TRUE creates a new database
conn <- dbConnect(SQLite(), "practicum2_p1.db")

customers = read_csv('customers.csv')
orders = read_csv('orders.csv')
router_info = read_csv('Router_Info.csv')
# copy does not function 
#copy_to(my_database,customers)

dbWriteTable(conn,"customers",customers,overwrite=TRUE)
dbWriteTable(conn,"orders",orders,overwrite=TRUE)
dbWriteTable(conn,"Router_Info",router_info,overwrite=TRUE)
```

## question 2
```{r}
str(customers)
```

```{r}
str(orders)
```


```{r}
str(router_info)
```

#### we can easily found these three tables have same quantity(499) of observation. I prefer to use full join to connect these tables for keeping all information. 

```{r}
dbGetQuery(conn," CREATE TABLE order_cus_router AS
           SELECT * 
           FROM orders 
           FULL JOIN customers ON orders.CustomerID = customers.CustomerID 
           FULL JOIN Router_Info ON orders.OrderID = Router_Info.OrderID")

```

#### After full join the number of observation still be same, which means there is no missing information at all.

## question 3
```{r}
cost_less_200 = dbGetQuery(conn, "SELECT *
           FROM orders
           WHERE Cost < 200
           LIMIT 20;")
kable(cost_less_200) %>%
  kable_styling("responsive", full_width = F)
```
## question 4
```{r}
full_info = dbGetQuery(conn,"SELECT * 
           FROM orders 
           FULL JOIN customers ON orders.CustomerID = customers.CustomerID 
           FULL JOIN Router_Info ON orders.OrderID = Router_Info.OrderID
           LIMIT 10;")
kable(full_info) %>%
  kable_styling("responsive", full_width = F)
```

## qeustion 5
```{r}
left_join = dbGetQuery(conn,"SELECT * 
           FROM orders 
           LEFT JOIN customers ON orders.CustomerID = customers.CustomerID
           LIMIT 10;")
kable(left_join) %>%
  kable_styling("responsive", full_width = F)
```

## question 6
```{r}
top_sum = dbGetQuery(conn,"SELECT State,
                    	   Year_Purchase,
                    	   count(*) as order_count,
                    	   sum(cost) as Cost
                 FROM order_cus_router
                 GROUP BY State, Year_Purchase
                 HAVING count(*) > 5
                 ORDER BY sum(cost) DESC;")
kable(top_sum) %>%
  kable_styling("responsive", full_width = F) %>%
  row_spec(1:3, bold = T, color = "white", background = "#D7261E")
```
## question7
Create a query that counts and distinctly counts
```{r}
count = dbGetQuery(conn, "select count(*) as totalCount,
                         count(distinct Year_Purchase) as YearDistinct
                    from order_cus_router")

kable(count) %>%
  kable_styling("responsive", full_width = F)
```

## question8
Create a query that finds the MIN, MAX, Sum, order by desc
```{r}
summary_query = dbGetQuery(conn, "select Year_Purchase,
                    	   min(cost) as minCost,
                    	   max(cost) as maxCost,
                    	   sum(cost) as totalCost
                    from order_cus_router
                    group by Year_Purchase
                    order by sum(cost) DESC")
kable(summary_query) %>%
  kable_styling("responsive", full_width = F)
```


## question9
Create a query that uses Between, and, <, >
```{r}
year_between = dbGetQuery(conn, "select orderID, customerID, Year_Purchase, Cost, firstname, lastname,
                    state, city
                    from order_cus_router
                    where Year_Purchase between 2018 and 2019 
                          and cost > 10 and cost < 30")
kable(year_between) %>%
  kable_styling("responsive", full_width = F)
```


## question10
Create a query that uses a wildcard, a join, and a group by clause
```{r}
wildcard_query = dbGetQuery(conn, "select orders.Year_Purchase,
                      		Router_Info.Reason,
                      		sum(cost) as totalCost
                    from orders
                    left join Router_Info on orders.OrderID = Router_Info.OrderID
                    where orders.SKU like '%BAS%'
                  group by orders.Year_Purchase,
                      		Router_Info.Reason;")
kable(wildcard_query) %>%
  kable_styling("responsive", full_width = F)

```

## question11
Create a query that finds the average and uses one of the following AND, OR, NOT
```{r}
ave_query = dbGetQuery(conn, "select Router_Info.Reason,
                      		AVG(cost) as averageCost
                    from orders
                    left join Router_Info on orders.OrderID = Router_Info.OrderID
                    where (orders.SKU like '%BAS%') or (orders.SKU like '%ADV%') 
                          and orders.Year_Purchase = 2021
                  group by Router_Info.Reason;")
 
kable(ave_query) %>%
  kable_styling("responsive", full_width = F)
```

##ending
```{r}
dbDisconnect(conn)
```

# part 2
## Question 1

```{r}
my_database = src_sqlite("practicum2_p2.db", create = TRUE) # create =TRUE creates a new database
conn <- dbConnect(SQLite(), "practicum2_p2.db")

country_info <- read_csv("country_info.csv")
dbWriteTable(conn,"country_info",country_info,overwrite=TRUE)

HIV_info <- read_csv("HIV_info.csv")
dbWriteTable(conn,"HIV_info",HIV_info,overwrite=TRUE)

dbGetQuery(conn,"CREATE TABLE HIV_country AS
           SELECT * 
           FROM country_info AS a
           FULL JOIN HIV_info AS b 
           ON a.Country = b.Entity and a.Year = b.Year")
```

#### From the the screenshot of these two tables, we can easily found these three tables have same quantity(499) of observation. I prefer to use inner join to connect these tables for explore the character of HIV.

## Question 2
```{r}
top_new_case = dbGetQuery(conn, "SELECT a.Country,
                         a.Year,
                         b.new_cases_of_hiv_infection,
                         a.School_enrollment_rate
                  FROM country_info as a
                  INNER JOIN HIV_info as b on a.Country = b.Entity and a.Year = b.Year
                  WHERE a.Country = 'North America'
                  ORDER BY b.new_cases_of_hiv_infection DESC;")
kable(top_new_case) %>%
  kable_styling("striped", full_width = F) %>% 
  row_spec(1:5, bold = T, color = "white", background = "#D7261E")

# inspect the structure of top_new_case
str(top_new_case)
# transform the School_enrollment_rate into num format for plot
top_new_case$School_enrollment_rate = sapply(top_new_case$School_enrollment_rate,as.numeric)
str(top_new_case)
ggplot(top_new_case) + 
  geom_point(aes(School_enrollment_rate,new_cases_of_hiv_infection)) +
  geom_smooth(aes(School_enrollment_rate,new_cases_of_hiv_infection)) +
  theme_bw()

ggplot(top_new_case) + 
  geom_point(aes(Year,new_cases_of_hiv_infection)) +
  geom_smooth(aes(Year,new_cases_of_hiv_infection)) +
  theme_bw()
```

#### From the plot above, it is hard to tell the relationship between new case of HIV and school_enrollment_rate. On the other hand, we should pay serious attention to the recent increase on the new case of HIV.

## question3
```{r}
living = dbGetQuery(conn, "select a.Country,
                   a.Year,
                   a.Country_Population,
                   1000000*b.number_of_people_living_with_HIV/cast(a.Country_Population as float) 
                   as Pct_of_people_living_with_HIV_per_Million,
                   1000000*b.deaths_total/cast(a.Country_Population as float) 
                   as Pct_of_deaths_from_HIV_per_Million
              from country_info as a
              inner join HIV_info as b on a.Country = b.Entity and a.Year = b.Year
              where a.Country = 'North America'
           ORDER BY Pct_of_people_living_with_HIV_per_Million DESC")

kable(living) %>%
  kable_styling("striped", full_width = F) %>% 
  row_spec(1:5, bold = T, color = "white", background = "#D7261E")

liv_plot = ggplot(living,aes(Pct_of_people_living_with_HIV_per_Million,Pct_of_deaths_from_HIV_per_Million, label = Year)) +
  geom_point() 
liv_plot + geom_smooth() +geom_text() + theme_bw()

```

#### From the line plot, we can find the living_rate for HIV patients is increasing from 2007 to 2015 and the Pct_of_deaths_from_HIV_per_Million is decreasing dramatically as the Pct_of_people_living_with_HIV_per_Million rocketing.

## question 4
```{r}
living_ratio = dbGetQuery(conn, "select a.Country,
                   sum(b.number_of_people_living_with_HIV/cast(a.Country_Population as float)) 
                   as living_people_with_HIV_divided_by_population
              from country_info as a
              inner join HIV_info as b on a.Country = b.Entity and a.Year = b.Year
              where a.Country <> 'North America'
              group by a.Country")
kable(living_ratio) %>%
  kable_styling("striped", full_width = F)
```

## question5
Produce 5 unique and interesting queries using both datasets.
```{r}
recent_year = dbGetQuery(conn, "select Entity,Year,
                    		number_of_people_infected_with_hiv,
                    		number_of_people_living_with_HIV, deaths_total 
                    from HIV_info
                    where year in (select max(year) from HIV_info)")
kable(recent_year) %>%
  kable_styling("responsive", full_width = F)

```


```{r}
query_2013 = dbGetQuery(conn, "select a.*,
                    		 b.*
                    from country_info as a
                    inner join HIV_info as b on 
                        a.Country = b.Entity 
                        and a.year = b.year 
                        and a.Year = 2013")
kable(query_2013) %>%
  kable_styling("responsive", full_width = F)
```



```{r}
death_total = dbGetQuery(conn, "select a.Entity,
              	   a.year,
              	   a.deaths_total,
              	   b.Country_Population,
              	   b.unemployment_rate,
              	   b.School_enrollment_rate
              from HIV_info as a
              inner join country_info as b on a.Entity = b.Country and a.year = b.Year
              where a.deaths_total > 10000")
kable(death_total) %>%
  kable_styling("responsive", full_width = F)
```

```{r}
max_newcase = dbGetQuery(conn, "select a.Entity,
                        a.Year, 
                        a.new_cases_of_hiv_infection
                from HIV_info as a
                inner join
                (select Entity, max(new_cases_of_hiv_infection) as newcases
                from HIV_info
                group by Entity) as b 
                    on a.Entity = b.Entity 
                    and a.new_cases_of_hiv_infection = b.newcases")
kable(max_newcase) %>%
  kable_styling("responsive", full_width = F)
```

```{r}
total_death = dbGetQuery(conn, "SELECT Entity,
                  SUM(new_cases_of_hiv_infection) as total_new_cases_of_hiv_infection,
                  SUM(deaths_total) as total_deaths
                  FROM HIV_info
                  GROUP by Entity;")
kable(total_death) %>%
  kable_styling("responsive", full_width = F)
```


```{r}
dbDisconnect(conn)
```

