---
title: "DA5020.A1.Xinhao.Sun.Rmd"
author: "Xinhao Sun"
date: "9/11/2022"
output: html_document
---

## Question 1
### 1)
```{r}
increment <- function(x, inc = 1) {
  x <- x + inc
  x
}
count <- 5
a <- increment(count, 2)
b <- increment(count)
count <- increment(count, 2)
```

#### I complete the data camp "Introduction to R". 
### 2)
```{r}
age = c("young", "middle", "old")
sex = c("female", "male", "male")

my_df = data.frame(age,sex)
my_df
```

#### A data frame has the variables of a dataset as columns and the observations as rows.(datacamp,2022)

## Question 2
```{r}
cars = c("Truck", "Car", "SUV")
mpg = c(11, 30, 24) 
cost = c(45000,25000,35000)

cars_df = data.frame(cars,mpg,cost)
cars_df


## a
### Select row 1 in column 3, what was selected? 
 cars_df[1,3]
### Select rows 1 through 3, what was selected? 
 cars_df[1:3,]
### Select the last column, what was selected?
 cars_df[,-1]

```
# Due the same names of variable, we need to clear the enviroment
```{r}
rm(list = ls())
```


## Question 3
```{r}
head(mtcars,3)
tail(mtcars,5)

```
## Question 4
```{r}

mtcars[,c("am","vs")] # categorical variables
mtcars[,c("mpg","hp")] # continuous variables
```
## Question 5
```{r error=TRUE}
ggplot(mtcars, aes(x=disp, y=mpg))
# for knit, we need to set error = TRUE
```
#### the chart is not displayed. we show use this code below to display 
```{r}
library(ggplot2) # ggplot is out-of-date
ggplot(mtcars, aes(x=disp, y=mpg))+geom_point()
```

## Question 6

```{r}

library(ggplot2)
str(mpg)

#use ggplot to draw scatter plot
ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = hwy, color = cyl))

```

#### In the scatterplot chart above, we can easily discover that the cars that have more cylinders tend to increasing engine displacement and to decreasing its highway miles per gallon.

## Question 7
```{r}
library(ggplot2)
#use ggplot to draw scatterplot
ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = hwy, color = cyl)) +
  facet_wrap(~drv)

```

#### In these scatterplot charts above, we can easily discover that the frontwheel drive cars are more likely to have less cylinders and to drive more miles in high-way, while the rearwheel drive cars tend to have more cylinders and to drive less miles in high-way. In addition, features of four-wheel drive cars are general when compared to frontwheel drive and rearwheel drive cars.
