---
title: "DA5020.A9.Xinhao.Sun"
author: "Xinhao Sun"
date: "11/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## initialization
```{r, message=FALSE}
rm(list=ls())
library(tidyverse)
library(forecast)
```

## Question 1
```{r}
url = "https://data.sfgov.org/api/views/rkru-6vcg/rows.csv?accessType=DOWNLOAD"
data = read_csv(url)
glimpse(data)
```

## Question 2
```{r}
data_Mar = data %>%
  filter(`GEO Summary` == "Domestic") %>%
  mutate(Date = as.Date(paste0(`Activity Period`, "01"), format="%Y%m%d")) %>%
  mutate(Year = format(Date, "%Y"), Month = format(Date, "%b")) %>%
  filter(Month == "Mar") %>%
  group_by(Year) %>%
  summarise(total_passengers = sum(`Passenger Count`))

  # as.Date must be three elements of date
  #mutate(Date = as.Date(`Activity Period`, format="%Y%m"))

glimpse(data_Mar)

#
ggplot(data_Mar, aes(x = Year, y = total_passengers, group = 1)) +
  geom_point()+
  geom_line() +
  labs( y = "Counts", title = "Total Passenger traveled in March each year.") +
  theme_bw()

```

##### From the line chart above, we can see the number of passenger in 2020 and 2021 drop significantly compared to 2019 since the pademic hited the U.S. , and government decided to lock down the whole country.

## Question 3
```{r}
# SMA
sma_19 = sum(data_Mar[11:13,2]) / 3
sma_error_19 = as.numeric(data_Mar[14,2]) - sma_19
sma_error_19
# for comparison, the rate should be more reasonable
sma_error_rate_19 = sma_error_19 / as.numeric(data_Mar[14,2])
sma_error_rate_19
```

#### Although the error of the prediction by SMA seems large with 64059.67 people, its error rate only 1.8%, which is acceptable.

## Question 4
```{r}
#WMA
weights_wma = c(3,5,7)
wma_19 = sum(data_Mar[11:13,2] * weights_wma) / sum(weights_wma)
wma_error_19 = as.numeric(data_Mar[14,2]) - wma_19
wma_error_19
# for comparison, the rate should be more reasonable
wma_error_rate_19 = wma_error_19 / as.numeric(data_Mar[14,2])
wma_error_rate_19
```

#### Comared with results of SMA, the results of WMA are more accurate both in quantity(17188.73) and rate(0.4%).

## Question 5
```{r}
# EWMA
#be careful
#3:10-3
#data_Mar[3:nrow(data_Mar)-3,]
#data_Mar[3:nrow(data_Mar)-4,]

pre_ses_data = data_Mar[3:(nrow(data_Mar)-4),]
ses_data = pre_ses_data[['total_passengers']]

# learn from overflow
ses_2019 = ses(ses_data,
    h = 1,
    alpha = 0.7,
    initial = "simple")$mean[1]
ses_error_19 = ses_2019 - as.numeric(data_Mar[14,2])
ses_error_19
# for comparison, the rate should be more reasonable
ses_error_rate_19 = ses_error_19 / as.numeric(data_Mar[14,2])
ses_error_rate_19
```

#### Although SES is more complicate than WMA, but its results is less accurate than that with 46005.37 and 1.3%.

## Question 6
```{r}
data_Mar <- data_Mar %>% mutate(Year = as.numeric(Year))
str(data_Mar)
lm_model <- lm(total_passengers ~ Year, data = data_Mar[3:13,])
lm_model

# according to the result above
predict_people = function(Year) 131769 * Year -262394559
predict_error = function(Year) {
  a = predict_people(Year)
  b = a - as.numeric(data_Mar[data_Mar$Year == Year,2])
  c = b / predict_people(Year)
  sprintf("The predicted values, error and error rate are %.2f, %.2f, %.4f",a,b,c)
}
predict_people(2019) 
predict_people(2020)

predict_error(2019)
predict_error(2020)
```
#### Compared with the prediction for 2019, the forecast for 2020 is unacceptable. But this is due to the black swan event, Covid-19 began to spread over the world.


## question 7
```{r}
ses_res = ses(ses_data,
                alpha = 0.7,
                initial = "simple")$fitted
#ses_res
mse_ses_data = pre_ses_data %>% 
  mutate(SE = (total_passengers - ses_res)^2) 

mse_ses_data
mse_ses = mse_ses_data %>%
  pull(SE) %>%
  mean()
mse_ses
```

```{r}

mse_lm_data = pre_ses_data %>% 
  mutate(lm_model_prediction = (lm_model$coeff[2] * data_Mar[3:13,1] + lm_model$coeff[1])$Year) %>%
  mutate(SE = (total_passengers - lm_model_prediction)^2)

mse_lm_data
mse_lm = mse_lm_data %>%
  pull(SE) %>%
  mean()
mse_lm

mse_ses > mse_lm
```

#### From the values above, it is obvious to find that the MSE for exponential model is larger than linear regression model for years from 2008 to 2018
